---
title: "3036274164_q2"
author: "PENG, Xin"
date: "2024-04-15"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
setwd("/Users/amypxx/Library/CloudStorage/OneDrive-TheUniversityofHongKong-Connect/adv stat/final exam")
```

# Q2 A
```{r}
library(metafor)
q2<-read.csv("exam_question2.csv")
head(q2)
summary(q2)

q2$se.prev <- (q2$prev.ub-q2$prev.lb)/(2*1.96)
q2.fe <- rma(yi=prev, sei=se.prev, slab=study, method="FE", data=q2)
q2.fe
```

# Q2 B
```{r}
summary(q2.fe)
Q <- q2.fe$QE
I2 <- (Q-(q2.fe$k-1))/Q * 100
I2
```
Yes, both Cochran’s Q test: p<0.0001 and I2=97.99% indicate substantial heterogeneity.

# Q2 C
```{r}
q2.re <- rma(yi=prev, sei=se.prev, slab=study, method="REML", 
               data=q2)

with(q2.re, c(b, ci.lb, ci.ub))
```
Based on a random effects model, the T2D prevalence among heavy drinkers is 17.58% (95%CI:12.87%~22.30%).

# Q2 D
```{r}
forest(q2.re, order=year, slab=paste(study, country,year, sep=", "), refline=0,  
       showweights=T, xlab="Prevalence, %",xlim=c(-3,3))
text(-3,22.5,"Study",pos=4,font=0)
text(3,22.5,"Prevalence [95% CI]",pos=2,font=0)
```

# Q2 E
```{r}
funnel(q2.re)

## Egger’s test
regtest(q2.re)
```
No obvious asymmetry can be seen from the funnel plot, and
there was no strong evidence of asymmetry from Egger’s test.
No evidence of publication bias.

# Q2 F
```{r}

prev_high <- c("Argentina", "Brazil", "Colombia", "Mexico", "Paraguay", "Peru")
high <- q2[q2$country%in%prev_high,]
non_high <- q2[!q2$country %in% prev_high,]

t_test <- t.test(high$prev, non_high$prev)
print(t_test)


q2$high <- 0
q2$high[q2$country%in% c("Argentina", "Brazil", "Colombia", "Mexico", "Paraguay", "Peru")] <- 1

q2.re.sen <- rma(yi=prev, sei=se.prev, slab=study, method="REML", data=q2, subset=(high==0))

# compare the original estimate vs estimate from sensitivity analysis
with(q2.re, c(b, ci.lb, ci.ub)) # original estimate for all regions

with(q2.re.sen, c(b, ci.lb, ci.ub)) # estimates for regions excluded Latin-America

```

# Q2 G

The estimated T2D prevalence among heavy drinkers is 17.58% (95%CI:12.87%~22.30%) based on the random effects model.
Sensitivity analysis shows that the inclusion of Latin-American regions where T2D prevalence is higher may impact the overall estimates.
The funnel plot showed that there was no evidence of publication bias.
