---
title: "3036274164_q1"
author: "PENG, Xin"
date: "2024-04-15"
output: word_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/amypxx/Library/CloudStorage/OneDrive-TheUniversityofHongKong-Connect/adv stat/final exam")
library(openxlsx)
library(ggplot2)
library(survival)
```

# Q1 A
```{r}
q1 <- read.csv('exam_question1.csv')
head(q1)
summary(q1)

lm1 <- glm(anxiety ~ covid + male + age + heart + comorbid, family = "binomial", data = q1)

ci <- exp(confint(lm1))
or <- exp(coef(lm1))

results <- cbind(
  Variable = rownames(ci),
  OR = round(or,2),
  round(ci,2)
)

knitr::kable(results, caption = "OR and 95% CI", digits = 2)
```
COVID-19 infections significantly increase the risk of anxiety disorder adjusting for various confounders.

# Q1 B
```{r q1b}
psq1 <- glm(covid ~ male + age + heart + comorbid, family = "binomial", data = q1)

summary(psq1)

q1$ps <- predict(psq1, type = "response")

p1 <- ggplot(q1, aes(x = ps, fill = as.factor(covid))) +
  geom_histogram(alpha = 0.5, position = "identity", bins = 30) +
  labs(title = "Distribution of Propensity Scores by COVID-19 Infection",
       x = "Propensity Score",
       fill = "COVID-19 Infection")

p1

# Exponentiate coefficients to get odds ratios
exp(coef(psq1))
```

Female, older people, people with heart disease and people with more comorbidity are more likely to get COVID-19 infection.

Common support in the range of 0.4 to 0.7.

# Q1 C

```{r q1c}

## stratification by propensity score quintiles

ps.boundary <- quantile(q1$ps, 0:5/5) 
ps.boundary
q1$psq <- cut(q1$ps, ps.boundary, right=F, include.lowest=T, label=1:5)
head(q1$psq)
head(q1)

q1$covid<-as.factor(q1$covid)

### age by PS strata
ggplot(q1, aes(x=covid, y=age)) + 
  geom_boxplot(aes(fill=covid)) + facet_grid(psq ~ ., labeller=label_both) + 
  coord_flip()


# sex by PS strata
ggplot(q1, aes(x=male)) + 
  geom_histogram(binwidth=1, fill='blue') + 
  facet_grid(covid ~ psq, labeller=label_both)

# heart disease by PS strata
ggplot(q1, aes(x=heart)) + 
  geom_histogram(binwidth=1, fill='blue') + 
  facet_grid(covid ~ psq, labeller=label_both)

# comorbidity by PS strata
ggplot(q1, aes(x=comorbid)) + 
  geom_histogram(binwidth=1, fill='blue') + 
  facet_grid(covid ~ psq, labeller=label_both)

```

Patient characteristics are more balanced within each propensity score strata.


# Q1 D

Conditional Independence: This assumption presumes that all known confounders are included in the model. However, the existence of unobserved confounders cannot be completely excluded.
Common Support: The overlap observed in the propensity score distribution between the treatment and control groups suggests that this assumption is reasonably met.

# Q1 E

```{r}
mean.cp <- aggregate(anxiety~covid+psq, data=q1, 
                     FUN=mean) # infection rates 
count.cp <- aggregate(anxiety~covid+psq, data=q1, 
                      FUN=length) # number of infections
cbind(mean.cp, count.cp$anxiety)

n.psq <- as.numeric(table(q1$psq))
n.psq 

covid.strata <- mean.cp$anxiety[mean.cp$covid==1]-
                 mean.cp$anxiety[mean.cp$covid==0] 

covid.est <- sum(covid.strata*n.psq)/sum(n.psq) 

## Estimated overall effect

round(covid.est,2) # Weighted Mean of the Stratum-specific effect

var.cp <- aggregate(anxiety~covid+psq, data=q1, FUN=var)
cbind(var.cp, count.cp$covid)

covid.var <- sum((var.cp$anxiety/count.cp$anxiety)*rep(n.psq, each=2)^2)/sum(n.psq)^2 

## Variance of estimated effect
covid.var 

## 95% CI
cl<-covid.est-1.96*sqrt(covid.var)
cu<-covid.est+1.96*sqrt(covid.var)

print(paste('95%CI:',round(cl,2),",",round(cu,2)))
```

# Q1 F

```{r}

model_ps <- glm(anxiety ~ covid + ps, data = q1, family = "binomial")
ci_ps <- exp(confint(model_ps))
or_ps <- exp(coef(model_ps))

results_table_ps <- cbind(
  "OR" = exp(coef(model_ps)),
  exp(confint(model_ps))
)

knitr::kable(results_table_ps, caption = "OR and 95% CI Using Propensity Score as a Covariate", digits = 2)
```


